---
title: "Inference high and low responders"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Se på filen "Inference high and low responders 2"


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Inference high and low responders}

# Bruk cellekjerner total RNA eller lignende

# Load needed packages
library(tidyverse)
library(readr)

# Download and save the file
download.file(url = "https://ndownloader.figstatic.com/files/14702420", destfile = "./data/hypertrophy.csv")

# Reading and attaching file to "hypertrophy"
hypertrophy <- read_csv("./data/hypertrophy.csv")

# View dataset
view(hypertrophy)

# Creating a hypothesis
## There is a statistical difference between high and low responders in 3RM squat

# Descriptive statistics: of the variable of interest (possibly including a figure)
## My variable of interest is 3 RM squat

### Choosing variables of interest (group and Squat_3RM_kg)
interest <- c("SUB_ID", "CLUSTER", "Squat_3RM_kg")

### Calculate summary statistics
hyp <- hypertrophy %>%
  
  select(all_of(interest)) %>%
  
  group_by(CLUSTER, Squat_3RM_kg) %>% 
  
  filter(!is.na(CLUSTER)) %>%          # Now the dataset only includes the two groups
  
  pivot_longer(names_to = "variable", 
               values_to = "values",
               cols = Squat_3RM_kg) %>%

  summarise(m = mean(values), 
            s = sd(values)) %>%
  
  print()   # Checking the tibble

# To see decimal numbers
data.frame(hyp)

view(hyp)   # Checking the table

### Making a boxplot ????????????
df <- data.frame(hyp = hyp) # Making a data frame

view(df)  # Checking the data frame

df %>%
  # Plots our samples on the y axis, and sets all x to 1.
  ggplot(aes(x = 1, hyp)) + 
  # Adds the boxplot
  geom_boxplot(width = 0.5) +
  # Scales the x axis to better plot the data
  scale_x_continuous(limits = c(0,2)) +
  # Set the name of the y axis
  labs(y = "Squat 3RM") +
  # The code below modifies the theme of the plot, removing 
  # the x axis
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(), 
        axis.line.x = element_blank()) 



# Inferential statistics: estimation of the effect and test of the hypothesis 
# Draw inference regarding the effect (is this effect true?)

hyp3 <- hypertrophy %>%
  select(SUB_ID, CLUSTER, T3T1__RNA) %>%
  filter(!is.na(CLUSTER)) %>%
  print()

t.test(T3T1__RNA  ~ CLUSTER, data = hyp3)

hyp3 %>%
  ggplot(aes(CLUSTER, T3T1__RNA)) + geom_boxplot()

## Tips til github
# If it doesn´t work with pushing the bottons, use te therminal instead. Type this in the terminal and push enter

# git status
# git pull





######## Creating a table of descriptive statistics ########
hypertrophy %>%
  
  # Selecting the variables of interest
  dplyr::select(SUB_ID, GROUP, CLUSTER, AGE, T1_BODY_MASS, 
                T1_TESTOSTERONE_ngdl, T3_TESTOSTERONE_ngdl, 
                T3T1__TESTOSTERONE_ngdl) %>% 
  
  pivot_longer(cols = AGE:T3T1__TESTOSTERONE_ngdl, 
               names_to = "Variable", 
               values_to = "values") %>% 
  
  # Removing NA from the data
  filter(!is.na(CLUSTER)) %>% 
  
  group_by(CLUSTER, values) %>% 
  
  # Calculating the mean and SD
  summarise(m = mean(values), 
            s = sd(values)) %>% 
  
  # Changing the names on columns and rows
  mutate(ms = paste(round(m,1), 
                    " (", 
                    round(s,1), 
                    ")"), 
        
        CLUSTER = factor(CLUSTER, 
                        levels = c("LOW", "HIGH"), 
                        labels = c("LOW (n = 10)", 
                                   "HIGH (n = 10)")),

         Variable = factor(Variable, 
                           levels = c("AGE", 
                                      "T1_BODY_MASS", 
                                      "T1_TESTOSTERONE_ngdl", 
                                      "T3_TESTOSTERONE_ngdl",
                                      "T3T1__TESTOSTERONE_ngdl"), 
                           labels = c("Age (years)", 
                                      "Body mass (kg)", 
                                      "Tetosterone T1 (ng/dl)", 
                                      "Tetosterone T3 (ng/dl)",
                                      "Change in tetosterone (ng/dl)"))) %>% 
  
  # Selecting the names and values of interest
  select(CLUSTER, Variable, ms) %>% 
  
  pivot_wider(names_from = CLUSTER, values_from = ms) %>%
  
  # Arrange the variables in the right order
  arrange(Variable) %>%
  
  # Making the format html and adding a caption
  kable(format = "html",
        caption = "Table 1 Descriptive statistics") %>% 
  
  # Adding a footnote
  footnote(general = "Values are Mean (SD)") %>%
  
  # Changing the table style and font
  kable_classic(full_width = F, html_font = "Times New Roman") %>%
  
  
  print()

# Hvordan få dette riktig? Trenger jeg det egentlig??


######################
#------------------------------------------ Erase, not part of the task
# Selecting the variables of interest 
hyp1 <- hypertrophy %>% # Attaching to "hyp1"
  select(SUB_ID, CLUSTER, T3_TESTOSTERONE_ngdl, T1_TESTOSTERONE_ngdl) %>% # Selecting variables of interest
  filter(!is.na(CLUSTER)) %>% # Removing participants that don´t belong to any group
  print()

# Using a paired, one sample t-test to see if there is an increase in tetosterone levels in HIGH and in LOW (using hyp1)
t.test(hyp1$T3_TESTOSTERONE_ngdl, hyp1$T1_TESTOSTERONE_ngdl, paired = TRUE, alternative = "two.sided")
#------------------------------------------

```

interest <- c("CLUSTER", "Squat_3RM_kg")

hypertrophy %>%
  
  select(all_of(interest)) %>%
  
  group_by(CLUSTER, Squat_3RM_kg) %>% 
  
  filter(!is.na(CLUSTER)) %>% print()




 
  summarise(m = mean(!is.na(CLUSTER)), 
            s = sd(!is.na(CLUSTER))) %>%
  
  print()
  
  
  hypertrophy %>%
  dplyr::select(CLUSTER, Squat_3RM_kg) %>%
  
  pivot_longer(cols = CLUSTER, Squat_3RM_kg, 
               names_to = "variable", 
               values_to = "values") %>%
  
  filter(!is.na(CLUSTER)) %>%
  
  group_by(CLUSTER, variable) %>%
  
  summarise(m = mean(values), 
            s = sd(values))